---
title: "hwk01_gapminder"
author: "Sarah"
date: "September 19, 2017"
output: 
  html_document: 
    keep_md: yes
---

#Loading packages 

```{r}
#install.packages("gapminder") #caught up at 10:30 in class, poor notes - re-read #instructions to follow
#install.packages("tidyverse")
library(gapminder)
library(tidyverse)
```

#Meet gapminder

```{r}
head(gapminder)
```

```{r}
tail(gapminder)
```

```{r}
str(gapminder)
```

```{r}
summary(gapminder)
```

```{r}
glimpse(gapminder)

```

```{r}
as_tibble(iris)
```

Ways to query basic info on a data frame:

```{r}
names(gapminder)
ncol(gapminder)
length(gapminder)
dim(gapminder)
nrow(gapminder)
```

Statistical overview
```{r}
summary(gapminder)
```


Also can visualize and place plots 
```{r}
plot(lifeExp ~ year, gapminder)

```

Look at the variables inside a data frame: 

```{r}
head(gapminder$lifeExp)
summary(gapminder$lifeExp)
hist(gapminder$lifeExp)
```

```{r}
#year is an integer variable but behaves like a categorical variable becasue there are so few unique values 
summary(gapminder$year)
table(gapminder$year)

```


```{r}
#truely categorical are country and continent 

class(gapminder$continent)
summary(gapminder$continent)
levels(gapminder$continent)
nlevels(gapminder$continent)

str(gapminder$continent) #R is really only storing integer codes 

```

Watch out for boobytraps! "I recommend you resolve to learn how to properly care and feed for factors. The pros far outweigh the cons. Specifically in modelling and figure-making, factors are anticipated and accommodated by the functions and packages you will want to exploit."


Count and display
-to quickly see that African countries are well represented in this dataset
```{r}
table(gapminder$continent)
barplot(table(gapminder$continent))
```

#ggplot2 play
```{r}
p <- ggplot(filter(gapminder, continent != "Oceania"),
            aes(x = gdpPercap, y = lifeExp)) # just initializes
p <- p + scale_x_log10() # log the x axis the right way
p + geom_point() # scatterplot
p + geom_point(aes(color = continent)) # map continent to color
p + geom_point(alpha = (1/3), size = 3) + geom_smooth(lwd = 3, se = FALSE)
p + geom_point(alpha = (1/3), size = 3) + facet_wrap(~ continent) +
  geom_smooth(lwd = 1.5, se = FALSE)
```

*pretty*

